---
title: "lab_02"
author: "derek willis"
date: "2023-01-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1** Create a new column that divides the expense amounts into the following categories - \$0-\$500, \$500-\$5000, \$5001-\$50000, \$50001\_\$100000, \$100001-\$500000 - giving each an appropriate text value, and count how many expenses are in each group. How would you describe the distribution of expenses in a sentence? Would you change the categories?

**A1.** A majority of expenses are less than $500.As the amount of money increases, the entries in that range decrease. There are 43 NA entries either because the amount was NA or because the amount was more than 500,000. I would add a category that was anything greater than 500,000 to include those entries. I would also separate the 0-500 category into two categories: 0-100 and 100-500 to give a more detailed description of the expenses.

```{r}
maryland_expenses_range <- maryland_expenses %>% 
  mutate(
    expense_range = case_when(
     amount<500 ~ "$0-$500",
     amount<5000 ~ "$500-$5000",
     amount<50000 ~ "$5001-$50000",
     amount<100000 ~ "$50001-$100000",
     amount<500000 ~ "$100001-$500000"
    )
  ) %>% 
  group_by(expense_range) %>% 
  summarise(count_expenses =n()) %>% 
  arrange(desc(count_expenses))
```

**Q2.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? 
**A2.** 27 departments have offered classes with Climate in the title. The department with the most classes is Atmospheric and Oceanic Science.

```{r}
umd_courses %>%
  filter(str_detect(title, "Climate")) %>% 
  group_by(department) %>% 
summarise(count_classes = n()) %>% 
  arrange(desc(count_classes))

```

**Q3.** Of the departments listed in A2, which one is the most interesting to you, and why?
**A3.** The most interesting department to me offering a class with Climate in the title is Germanic Studies because I did not expect a foreign language major to appear on the list.

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name). Write a sentence describing the pattern of spending over time.
**A4.** As the month grows closer to the election, the more spending increases. Spending is greatest in October 2022, the month before the election and in June 2022, the month before the Maryland primaries. The third highest month is July 2022 the month of the primaries. 

```{r}
maryland_expenses_date <- maryland_expenses  %>%
  filter(committee_name == "Moore  Wes For Maryland") %>% 
  mutate(expenditure_date=mdy(expenditure_date)) %>% 
mutate(month = floor_date(expenditure_date, "month")) %>% 
    group_by(month) %>% 
    summarise(total_amount = sum(amount)) %>% 
    arrange(desc(total_amount))
  

```

**Q5.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab.
**A5.** My question was what was the total amount spent on online advertising by campaigns. Between all campaigns in Maryland, 5,614,719 was spent on online advertising.

```{r}
maryland_expenses_online_ads <- maryland_expenses %>%
  filter(expense_purpose == "Online Advertising") %>% 
summarise(total_amount = sum(amount))
```
