---
title: "Reverse Engineering Project (North Carolina In-Session Campaign Donations"
author: "Olivia Janik and Mitchell Hang"
date: "March 4, 2023"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [CEOs gave heavily during legislative session, exposing loophole in NC's fundraising ban](https://www.wral.com/in-session-campaign-donations/17716104/)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
options(scipen = 999)
library(tidyverse)
library(lubridate)
library(janitor)
library(stringr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data
NC_donors <- read_csv("data/donors0103-0701.csv")

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

cleaned_NC_donors <- NC_donors %>%
  clean_names() %>%
  mutate(date=mdy(date)) %>%
  mutate(amount=parse_number(amount)) %>%
  mutate(zip5 = str_sub(zip, start = 1L, end = 5L)) %>%
  mutate(zip5 = case_when(
    zip5 == "Vario" ~ as.character(NA),
    zip5 == "UB7 O" ~ as.character(NA),
    zip5 == "UB7 " ~ as.character(NA),
    TRUE ~ zip5
  ))
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “Senate President Pro Tem Phil Berger's campaign alone raised nearly \$222,000 during session. Twenty-three people gave him the maximum allowed: \$5,200.”

* **Analysis summary**: Based on the dates provided for the North Carolina session, we can confirm that 23 people donated \$5,200 (which was indeed the maximum amount given during this time) to Berger's campaign. When we calculated all of the amounts together, the total amount we got was \$221,925, which did match with the article.

```{r}
# Put code to reverse engineer sentence here
NC_donors_1 <- cleaned_NC_donors %>%
 filter(between(date, as.Date('2018-05-16'), as.Date('2018-06-29'))) %>%
filter(str_detect(committee_name, "BERGER")) %>%
  group_by(`name_clean`) %>%
  summarise(total_amount = sum(amount)) %>%
  arrange(desc(total_amount)) %>%
  bind_rows(summarize_all(., ~if(is.numeric(.)) sum(.) else "Total"))

# Display results of code below this codeblock
view(NC_donors_1)
```

### Sentence 2

* **Sentence text**: "This year, the week saw more than $1 million given to state legislative campaigns, the state Republican and Democratic parties and the governor's campaign account."

* **Analysis summary**: Using this code, we were able to confirm that more than /$1 million was donated to state legislative campaigns, the Republican and Democratic parties and the governor's campaign. The total was 1,052,718.4 dollars.

```{r}
# Put code to reverse engineer sentence here
NC_donors_2 <- cleaned_NC_donors %>% 
filter(between(date, as.Date('2018-05-09'), as.Date('2018-05-16'))) %>% 
  group_by(source) %>%
  summarise(total_amount = sum(amount)) %>%
  adorn_totals
  

# Display results of code below this codeblock
view(NC_donors_2)
```

### Sentence 3

* **Sentence text**: “There was a major bump during the first week of session this year for the state parties, but that's because three large donors – LinkedIn co-founder Reid Hoffman, Durham investor Greg Lindberg and Lindberg's Eli Research company – donated \$1.3 million that week to the state Republican and Democrat parties.”

* **Analysis summary**: We gathered the party donations during the first week of session in 2018 (May 16th to 23rd) and the total amount that we calculated from the three names "GREG LINDBERG," "REID GARRETT HOFFMAN" and "ELI RESEARCH LLC" was \$1,250,000. So the total amount during that week wasn't necessarily \$1.3 million, but we guess many publications prefer rounding up to the nearest tenth decimal when using figures like these.

```{r}
# Put code to reverse engineer sentence here
NC_donors_3 <- cleaned_NC_donors %>%
  filter(between(date, as.Date('2018-05-16'), as.Date('2018-05-23'))) %>%
  filter(str_detect(source, "parties")) %>%
  group_by(name_clean) %>%
  summarize(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
# Display results of code below this codeblock
view(NC_donors_3)
```

### Sentence 4

* **Sentence text**: “This year, 18 of the top 20 donors to legislators during the week before session were PACs. The other two were Strobel and Delmer.”
* **Analysis summary**: We were not able to confirm this finding because when we filtered for donations that came in the week before session, we found that only 14 of the top 20 donors were PACS. The other six were Greg Lindberg, George A Luecke, George Lewis Strobel II, Robin E Delmar, Matthew S Cullinan and Temple Sloan. 

```{r}
# Put code to reverse engineer sentence here
NC_donors_4 <- cleaned_NC_donors %>%
 filter(between(date, as.Date('2018-05-09'), as.Date('2018-05-15'))) %>%
  group_by(name_clean) %>%
  summarize(total_amount = sum(amount)) %>%
  arrange(desc(total_amount))
# Display results of code below this codeblock
view
```

### Sentence 5

* **Sentence text**: “Going back to the beginning of the year, people connected to Prestage Farms gave Moore \$31,000 in campaign donations. Half are dated two-and-a-half months before the session began.”
* **Analysis summary**: People associated with Prestage Farms gave Moore's committee $31,200 across six payments in 2018. Three came on March 1, 2018, which is about two and a half months before the session began. The other three donations occurred on June 30, 2018.

```{r}
# Put code to reverse engineer sentence here
NC_donors_5 <- cleaned_NC_donors %>%
  filter(str_detect(employer_name, "PRESTAGE FARMS") & str_detect(committee_name, "MOORE")) %>% adorn_totals
  
# Display results of code below this codeblock
```

-30-

