---
title: "lab_06"
author: "derek willis"
date: "2022-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(lubridate)
library(janitor)

# load your libraries

```

## Getting Started

This week we'll be combining and joining data to make it more useful and to ask some questions about it. We'll use some of our 911 overdose call data to do this. The first thing we want to do is to combine multiple counties' data into a single dataframe so that we can ask some questions. First, let's combine data from Cecil, Carroll and Allegany counties into a new dataframe.

*Before* you combine them you'll need to clean up some of the column names to make the data easier to work with - make sure you have loaded the library to do that. You also need to make sure that each column has the same name and datatype (you can check the datatypes using `glimpse`). If any of them does not, you need to fix that.

```{r}
# load and combine the call data from those three counties
allegany_911 <- read_csv("data/allegany_911.csv")
cecil_911 <- read_csv("data/cecil_911.csv")
carroll_911 <- read_csv("data/carroll_911.csv")
maryland_zcta <- read_csv("data/maryland_zcta.csv")
baltimore_911 <- read_csv("data/baltimore_911.csv") 
```
```{r}
allegany_911 <- allegany_911 %>% 
  clean_names() %>% 
  mutate(location = str_to_lower(location), county = str_to_lower(county))
  
carroll_911 <- carroll_911 %>% 
  clean_names() %>%
  mutate(location = str_to_lower(location), county = str_to_lower(county))
cecil_911 <- cecil_911 %>% 
  clean_names() %>%
  mutate(location = str_to_lower(location), county = str_to_lower(county))

combined_911 <- bind_rows(list(cecil_911, allegany_911, carroll_911))


```

Then we'll join some demographic data with Baltimore City 911 calls and ask some questions of that. For that, you'll need to load the Baltimore City 911 data and the ZCTA demographic data, and join them based on the zip code. Remember, the zip codes in both dataframes must be the same datatype (and should be characters). You should start with your Baltimore 911 dataframe in creating your join. 

```{r}
# load and join the Baltimore-related data
baltimore_911_with_demographics <-baltimore_911 %>% 
  left_join(maryland_zcta, by=c("zip_code"= "ZCTA5N")) %>%
  clean_names()
```

## Answer questions

Q1. Let's start with our data from Allegany, Cecil and Carroll counties. Write code to generate a dataframe that shows the total number of calls for each county. What's the order of counties from most calls to least?

A1. Carroll County had the most 911 overdose calls with 448, then Cecil County, then Allegany. There was not a huge difference between the three counties in the number of calls. Carroll only had 10 more calls than Cecil and 30 more than Allegany.

```{r}
combined_911_A1 <- combined_911 %>%
  group_by(county) %>%
summarise(count_calls = n()) %>%
  arrange(desc(count_calls))
```

Q2. In your combined dataframe that you used in A1, add a column for the month of each call, then write code to show the total number of calls per county and month. Are there any outliers or noteworthy results? Describe the general pattern of the results. 

Then calculate the most calls per county and location. Which location had the most calls - where and what is that location?

A2. The month with the most callls is August with 149 calls. The month with the least calls is January with 88 calls. While there is some difference in the amount of calls each month. There are no true outliers.

Of the three counties, the address with the most calls is 314 grove neck rd Earleville, MD with 15 calls. After googling the address I discovered this location is Recovery Centers of America at Bracebridge Hall, an inpatient alcohol and drug addiction treatment center. It has more than twice the amount of calls than the next locations on the list. In Caroll County, the address with the most calls is 10 George Street, a townhouse in Westminster with 7 calls. In Allegany County, 10 Taylor St. had the four calls, the most in the county. This address is a single family home in Frostburg.

```{r}
combined_911_A2 <- combined_911 %>%
  mutate(month=month(date, label=TRUE, abbr=FALSE)) %>%
  group_by(county, location) %>%
summarise(total_calls= n()) %>%
  arrange(desc(total_calls))
```

Q3. Let's turn to the Baltimore City data. You're interested in exploring calls in zip codes with a high percentage of children under 18. Write code to isolate calls in zip codes where the percentage of under 18 population is at least 75% and then show the zip code and population and how many calls occurred in each of those zip codes. 

Where are those zip codes in the city?

Add a column to your code that calculates the number of calls per 1,000 people for those zip codes. To calculate a per capita rate, you might find this short guide handy: https://observablehq.com/@palewire/per-capita-calculator.

Which zip code has the highest rate of calls per 1,000 people? Find the neighborhoods in that zip code that are listed in the data - you can use group_by or distinct to do this. What are some of those neighborhoods, and what else can you tell me about the population there?

A3. The zip code with the highest calls per 1000 residents is 21201, a zip code in downtown Baltimore. In this zip code, 52.4 percent of the population is Black and 32.3 percent is white. The median age is 30.3 years old and 81.5 percent of residents are under 18. 90.6 percent of residents rent.

```{r}
baltimore_911_with_demographics_A3 <- baltimore_911_with_demographics %>%
filter(pct_under_18 > 75) %>%
  group_by(zip_code, population) %>%
  summarise(total_calls=n()) %>%
  mutate(per_capita = (total_calls/population) * 1000)
```

Q4. Choose your own adventure: I want you to come up with a question you could ask of the Baltimore City data that combines both the calls and the demographic information. You must use both demographics and the overdose call data in your work.

A4: In zip codes with the most drug overdose calls, what is the home ownership percentage and what percent of residents rent? In the zip code with the most calls (774), 70 percent of residents are renters and 21.6 percent own. Zip codes where more than 50 percent of residents own did not have more than 100 overdose calls, while zip codes with 70 or more percent renters each had 300 or more calls.
```{r}
baltimore_911_with_demographics <- baltimore_911_with_demographics %>%
  group_by(zip_code, pct_homeowner, pct_rent) %>%
  summarise(total_calls=n()) %>%
  arrange(desc(total_calls))
  
```
