---
title: "lab_05"
author: "derek willis"
date: "2023-03-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

- Tabula

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(janitor)
library(lubridate)
```


## Get Our PDF

We'll be working with the [911 overdose calls from Baltimore County](https://drive.google.com/file/d/1qkYuojGF_6WKFr5aNQxmewDzcKyOiJFr/view?usp=share_link). You'll want to download it to a place you'll remember (like your Downloads folder, or the labs folder in your repository). The goal is to extract the tables within it, export that to a CSV file, load it into RStudio and ask some questions.

## Extract Data from PDF Using Tabula

Start Tabula, then go to http://127.0.0.1:8080/ in your browser. Click the "Browse" button and find the PDF file and click "open", and then click the "Import button" in Tabula. This will take a few seconds or longer.

This PDF has a single table spread over multiple pages to extract. We're going to make a single dataframe from this table, exporting it to a CSV file that you will load into R. In Tabula, highlight the table and click the "Preview & Export Extracted Data" button. You may want to play with including or excluding the column headers - YOU SHOULD HAVE FIVE COLUMNS OF DATA.

Save the CSV (it should be called `tabula-Baltimore County; Carey, Samantha log OD.csv` by default) to your lab_05/data folder.

From there, you will need to read in the data, and add or fix headers if necessary. You can choose to include the headers from the PDF in your exported CSV files OR to exclude them and add them when importing. `read_csv` allows us to do this ([and more](https://readr.tidyverse.org/reference/read_delim.html)).

## Load and clean up the data in R

You will need to read in and clean up the data so that it can be used for analysis. By "clean" I mean the column headers should not contain spaces and they should have meaningful names, not "x1" or something similar. How you do that is up to you, but you can use select() with or without the minus sign to include or exclude certain columns. You also can use the `rename` function to, well, rename columns. Importantly, you'll need to ensure that any columns containing a date actually have a date datatype. Our friend `lubridate` can help with this.

```{r}
baltimorecounty_911overdoses <- read_csv("data/tabula-baltimorecounty-911overdoses.csv", col_names = FALSE) %>% 
  clean_names() %>%
  rename(date = x1, time = x2, case_number = x3, event_type = x4, location = x5) %>%
  mutate(date=mdy(date))

```


## Answer questions

Q1. Write code to generate the number of calls that occurred on each date. Which date in 2022 had the most overdose calls, and how many? Look at the total number of rows in your result and explore the range of dates - based on your result, do you believe there are any days with no overdose calls at all? Explain why or why not.

A1. In 2022, the day with the most 911 overdose calls was July 14, 2022 and October 4, 2022, both with 23 calls. According to my code, there were 329 entries for dates, which means that there were 0 calls on 36 days in 2022. However, there is a chance that calls happened on those 36 days but were not recorded in this data set. The fact that there are no entries from January 2022 makes me hesitant.

```{r}
baltimorecounty_A1 <- baltimorecounty_911overdoses %>%
  filter(between(date, as.Date('2022-01-01'), as.Date('2022-12-31'))) %>%
group_by(date) %>%
  summarise(count_date = n()) %>%
  arrange(desc(count_date))
```

Q2. You want to understand if there's a pattern in the day of the week that overdose calls are made. Add a column to your dataframe that displays what day of the week each date represents. You should search for how to do that using lubridate. Then write code to calculate the number of calls for each day of the week, and add a column to that result that calculates the percentage of all calls that occurred on each day of the week (so you want a dataframe with the day of the week, total number of calls and the percentage of calls on that day out of the total number of all calls). Describe your findings to me.

A2. The day of the week with the most calls is Saturday, with 15.5% of all calls. Sunday and Friday also have about 15% of the calls. Thursday has the lowest percentage with 12.8%. It is not surprising to me that Friday and Saturday have higher numbers because people are more likely to party and partake in illicit substances when they don't have to work the next day. That is why I am surprised that Sunday has the second highest percentage of calls.

```{r}
baltimorecounty_A2 <- baltimorecounty_911overdoses %>%
mutate(weekday = wday(date, label=TRUE, abbr=FALSE)) %>% 
group_by(weekday) %>%
summarise(count_weekday = n()) %>%
 mutate(percent_total = count_weekday/4112)

```

Q3. Now let's look at locations. Which ones have the most calls? How would you describe them (feel free to search for more information on them)? Is there anything about the structure of the original data that might make you less confident in the counts by location or date?

A3. An estimated 36 overdose calls came from 4540 Silver Spring Road. According to google maps and real estate websites this location appears to be the entrance to a trail off the main road. The locations with the second most calls is the Baltimore County Police Department, Precinct 2 and Baltimore County Police Department, Precinct 6 with 13 calls each. Between call locations at actual police precincts and the lack of calls in January 2022, I am hesitant about some pieces of this data.
```{r}
baltimorecounty_A3 <- baltimorecounty_911overdoses %>%
group_by(location) %>%
summarise(count_location = n()) %>%
arrange(desc(count_location))

```

Q4. What's the best story idea or question you've seen as a result of the work you've done in this lab?

A4. I think the 4540 Silver Spring Road property is worth looking into for multiple reasons. I want to know how often it is that one location has more than 30 overdoes calls in a little over a year. According to Zillow, this property was sold for $100,000 on January 18, 2023, after all of these overdose calls occurred. From what I can tell on Google maps, the trail at this address leads toward Indian Rock Park and Perry Hall High School. Was this property bought by the government or maybe the school in an effort to make the area safer for students? If not, did the buyer know about the 36 drug overdose calls from this address? Is it a common occurance for drug overdose calls to happen near high school?
