---
title: "lab_01"
author: "Derek Willis"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

**Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first.
**A1** Engineering, Cooperative Education

```{r}
umd_courses %>%
  group_by(department) %>% 
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats),
  ) %>% 
  arrange(desc(mean_seats))
  
```

**Q2** Find out the greatest number of times any individual class has been offered in our dataset. You must show the greatest number of classes first.
**A2** Doctoral Dissertation Research has the most courses listed with 1229 classes.

```{r}
umd_courses %>%
  group_by(title) %>% 
  summarise(
    count_classes = n()
  ) %>% 
  arrange(desc(count_classes))
```

**Q3** Count the number of courses for each instructor and department. Which instructor has taught the most? 
**A3** Keith Yearwood in Geographic Sciences has taught 58 classes in this data period.

```{r}
umd_courses %>% 
group_by(instructors, department) %>% 
  summarise(
    count_classes = n()
  ) %>% 
  arrange(desc(count_classes))
```

**Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know _how_ to accomplish this; I want you to describe in words what would make the answer easier to see.
**A4**If I was able to exclude NA and Instructor:TBA listings from the data, I would be able to see the instructors with the most courses easier. It is also difficult to know if there are instructors with the same names being listed or instructors teaching in multiple departments. Some classes are also listed as graduate and undergraduate classes but are taught in the same section.

**Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting?
**A5** The question was which department offers the most classes per term. After running a code that grouped by department and term, I discovered that the Business and Management department offered the most courses for the fall and spring terms each semester. Because of the way I grouped and arranged the data, with the most courses listed highest, it is more difficult to find what department had the most winter and summer term classes.

```{r}
umd_courses %>% 
  group_by(department, term) %>% 
summarise(
  count_classes = n()
) %>% 
  arrange(desc(count_classes))
```
