---
title: "Spotify and YouTube Music Data Analysis Project"
author: "Mitchell Hang & Olivia Janik"
date: "4/14/23"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are analyzing a dataset which features statistics for the top 10 songs of various artists on Spotify and their corresponding YouTube videos. (https://www.kaggle.com/datasets/salvatorerastelli/spotify-and-youtube)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(ggthemes)
library(scales)
library(corrr)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 
The Spotify and Youtube data has 20,718 observations of 28 variables. The main variables we will be working with in this data are number of streams, danceability, valence, key, and views on Youtube. After looking more closely at the data, we believe that we will be able to answer all of the questions we posed in the previous milestone. There were 1,655 songs with NA in at least one of their columns so we removed them from the data set for simplification.

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

Spotify_Youtube <- read_csv("data/Spotify_Youtube.csv")

Spotify_Youtube_clean <- Spotify_Youtube %>%
  clean_names() %>%
  drop_na()


```

## Questions to Analyze
In this notebook, we are using data analysis to answer five newsworthy questions.

### Question 1

* **Question text**: What is the correlation between a song’s “danceability” and its number of streams?
* **Analysis summary**: In this question we separated songs into 10 danceability intervals and coded the average mean for each interval. Through this code, we found that songs with higher danceability tend to have more streams on average. Songs with a danceability of less than 0.2 have far lower average number of streams while the danceability intervals with the most number of streams are the top 3 intervals. We also created a scatter plot of songs danceability vs. number of streams with a trendline, while the trendline showed there was a positive relationship between the 2 variables, it did not appear to be significant based on its slope. To confirm this, we ran a correlate test, which showed that the r-value is 0.072, which means there is a very weak positive relationship between the two.

```{r}
Spotify_Youtube_clean_Q1 <- Spotify_Youtube_clean %>%
  mutate(stream_in_millions = (stream/1000000)) %>%
    mutate(danceability_interval = case_when(
    danceability <= 0.1 ~ "0-0.1",
    danceability <= 0.2 ~ "0.101-0.2",
    danceability <= 0.3 ~ "0.201-0.3",
    danceability <= 0.4 ~ "0.301-0.4",
    danceability <= 0.5 ~ "0.401-0.5",
    danceability <= 0.6 ~ "0.501-0.6",
    danceability <= 0.7 ~ "0.601-0.7",
    danceability <= 0.8 ~ "0.701-0.8",
    danceability <= 0.9 ~ "0.801-0.9",
    danceability <= 1 ~ "0.901-1"
    )) %>%
   group_by(danceability_interval) %>%
  summarise(average = mean(stream_in_millions)) %>%
  arrange(desc(average))
  
Spotify_Youtube_clean_Q1 %>%
    ggplot() +
  geom_bar(aes(x=danceability_interval, weight=average)) +
  theme_few () +
  scale_fill_few("Light") +
    labs(
    title="Songs with higher danceability score tend to have more streams",
    x = "Danceability interval",
    y = "average number of streams in millions",
    caption = "source: Kaggle"
    
  )
Spotify_Youtube_clean %>%
  ggplot() +
  geom_point(aes(x=danceability,y=stream)) +
  geom_smooth(aes(x=danceability,y=stream ), method="lm")

Spotify_Youtube_clean %>%
  select(danceability, stream) %>%
  correlate()
  
#glimpse(Spotify_Youtube_clean)
# Display results of code below this codeblock

```

### Question 2

* **Question text**: What mood of song gets more streams, happy songs (0.8 valence or higher) or sad songs (0.2 valence or lower)?
* **Analysis summary**: Songs with a valence of 0.8 or more (happy songs) have an average 144.9 million streams, more than songs with a valence of 0.2 or less, which have an average of 125.3 million streams. However, the valence interval with the lowest average streams is 0.4 to 0.6, songs with a medium valence. 

```{r}
# Put code to reverse engineer sentence here
Spotify_Youtube_clean_Q2 <- Spotify_Youtube_clean %>%
  mutate(stream_in_millions = (stream/1000000)) %>%
    mutate(valence_interval = case_when(
    valence <= 0.2 ~ "0-0.2",
    valence <= 0.4 ~ "0.2-0.4",
    valence <= 0.6 ~ "0.4-0.6",
    valence <= 0.8 ~ "0.6-0.8",
    valence <= 1 ~ "0.8-1",
    )) %>%
   group_by(valence_interval) %>%
  summarise(average = mean(stream_in_millions)) %>%
  arrange(desc(average))
# Display results of code below this codeblock

```

### Question 3

* **Question text**: What is the correlation between a song’s key and its valence? Do certain keys register positive emotions while others register negative emotions?
* **Analysis summary**: There seems to be no apparent relation between a song's key and its valance. Initially, we thought that a natural key would result in a positive emotion being registered whereas an accidental (sharp or flat) would register a negative emotion, but upon graphing the respective data points, we notice that each of the keys were filled for all possible amounts of valence from 0 to 1.


```{r}
potify_Youtube_clean_Q3 <- Spotify_Youtube_clean %>%
 ggplot() +
   geom_point(aes(x=key, y=valence)) +
    theme_minimal() +
    labs(
    title="No apparent relationship between song's key and valence",
    x = "Key (0 is C, 1 is C#, 2 is D, etc.)",
    y = "Valence",
    caption = "Source: Kaggle"
  )

```

### Question 4

* **Question text**: What artist has the most songs in the top 100 most streamed songs on Spotify?
* **Analysis summary**: The artist who has the most songs in the top 100 most streamed songs on Spotify is Post Malone, with five songs: "rockstar (feat. 21 Savage)," "Sunflower - Spider-Man: Into the Spider-Verse," "Circles," "Better Now" and "Congratulations."

```{r}
# Put code to reverse engineer sentence here

Spotify_Youtube_clean %>%
  arrange(desc(stream)) %>%
  slice(1:100) %>%
  group_by(artist) %>%
  summarize(total_songs = n()) %>%
  arrange(desc(total_songs))

Spotify_Youtube_clean %>%
  group_by(artist) %>%
  summarise(average = mean(stream)) %>%
  arrange(desc(average))

# Display results of code below this codeblock

```

### Question 5

* **Question text**: Which song has the largest difference between Spotify streams and Youtube views from its official video. 
* **Analysis summary**: "Despacito" by Luis Fonsi feat. Daddy Yankee has the largest difference between Spotify streams and YouTube views, with a difference of more than 6.5 billion, nearly 2 billion more than the difference of the next song, Wheels on the Bus by CoComelon. The official video for "Despacito" was once the most viewed YouTube video ever, so it makes sense as to why the difference in streams and views is so large. The only artist to have two songs in the top 10 largest differences was CoComelon, a children's music YouTube channel. Wheels on the Bus by CoComelon has a higher views-to-stream ratio than Despacito, most likely because parents play the video to entertain their children rather than the song.

```{r}
# Put code to reverse engineer sentence here

Spotify_Youtube_clean %>%
  filter(official_video = TRUE) %>%
  mutate(difference = abs(views - stream)) %>%
  mutate(views_to_stream_ratio = views/stream) %>%
  arrange(desc(difference)) %>%
  summarize(artist, track, views, stream, difference, views_to_stream_ratio)

# Display results of code below this codeblock

```

-30-